Can historical data predict population responses to climate change experiments?

### 2016-12-14 precip/manuscript/precip_ms.tex
adding edits to the abstract

\title{Can historical data predict ecological responses to novel climate conditions?}

@@

\begin{enumerate}
\item Climate is thought to be an important driver of population ecology however it can be difficult to determine exactly which aspects of climate affect populations and at which stages in an organisms life-cycle.  Moreover, there have been few tests of whether statistical models fit to historical climate and population data can accurately predict the future response of populations to natural and experimentally imposed climate conditions.   
\item Using long-term observational data collected from the 1920's to the 1950's from a sagebrush steppe plant community, previous research has reported a wide variety of climate effects on four dominant plant species. We tested whether statistical models fit to these historical observational data could predict the population response of these plant populations to natural and experimentally imposed climate variation some 60 years after the historical data collection ended. 
\\item We established sixteen new plots in the same location as the historical data and assigned each to either drought or irrigation treatments imposed using rainfall shelters and automatic sprinklers. The original plots were also monitored and served as a control. We fit statistical models for survival, growth and recruitment of each species as functions of seasonal temperature and soil moisture in the historical data.  We then used these models to predict the treatment effects observed in contemporary data.  
\item  For   \textit{Poa secunda} recruitment, the removal treatment was actually negative, indicating that our baseline model overestimated competitive release.
\item For growth, the removal treatment effect was significant and positive for two species, \textit{Poa secunda} and \textit{Pseudoroegneria spicata}, indicating that the baseline model underestimated competitive release. However, including information about the location of these grass individuals with respect to removed \textit{Artemisia} failed to improve the growth regressions, raising questions about the mechanisms driving the positive response to removal.
\item For three of the four species, individual based models and integral projections models based on the vital rate regressions showed that removal treatment effects had little impact on population growth rates or simulated equilibrium cover following a species removal. For \textit{P. spicata}, the population models that included effects of \textit{Artemisia} removal projected greater competitive release than the baseline models. 
\item \emph{Synthesis}: Although models based on observational data did not perfectly predict species responses to removal, overall the results increase our confidence that competitive release is in fact small for at least three of our four study species. The implication is that interspecific resource competition is not the primary factor determining the dynamics of this community.

>>>>

\title{Can historical data predict population responses to climate change experiments?}

@@

\begin{enumerate}
\item Climate is thought to be an important driver of population ecology however it can be difficult to determine which aspects of climate affect populations and at which stages in an organisms life-cycle.  Moreover, there have been few tests of whether statistical models fit to observational data on climate and populations can accurately predict the future responses of populations to natural or experimentally imposed climate conditions.   
\item Using observational data collected from the 1920's to the 1950's from a sagebrush steppe plant community, previous research has reported a wide variety of climate effects on four dominant plant species. We tested whether statistical models based on these historical data could predict how these species respond to natural and experimentally imposed climate variation from 2006 to 2016, some 60 years after the historical data collection ended. 
\\item We established sixteen new plots in the same location as the historical data and assigned each to either drought or irrigation treatments imposed using rainfall shelters and automatic sprinklers. The original plots were also monitored and served as ambient climate controls. Using the historical data, we fit statistical models for survival, growth and recruitment of each species in response to seasonal temperature and soil moisture.  We used these models to predict variation in species vital rates in the experiment and used a simple individual based population model to predict one year ahead population growth rates in each of the experimental plots.   
\item Over the course of the experiment, average plot cover of \textit{Hesperostipa comata} and \textit{Pseudoroegneria spicata} declined, while cover of \textit{Poa secunda} showed a trend towards increase in the irrigated plots. At the level of individual vital rates, experimental drought led to decreases in \textit{} growth, survival, while irrigation led to increases in \textit{}. 
\item Vital rate models fit to the historical data correctly predicted significant increases in X and X and significant decreaes in X and X.  However, the historical models predicted significant changes in X X and X that were not observed in the experiment.  Moreover, the experiment led to a significant change in X that was not predicted by the historical data. 
\item Our individual based population model more accurately predicted changes in the population growth rates of \textit{} in the experimental plots than a model without climate effects.  However for the other X species predictions made from the climate model were less accurate than models without climate.   
\item \emph{Synthesis}: Out of all four species examined, climate covariates only appeared to increase our predictive accuracy for one species: \textit{Hesperostipa comata}. In general the relatively weak responses to the experimental treatments agrees with the low explanatory power of climate covariates in explaining the vital rates and population dynamics of these species in this system. The highly variable environment and water limited conditions may select for species with a conservative non-responsive strategy to variation in climate both natural and experimentally imposed.
